### 2020-09-04 supplementary.Rmd
add supplementary material

---
title: "Supplementary Materials"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  error = FALSE)
```

# Intro

Here we provide side by side view of animations provided in the case studies of our paper:

>  Burning sage: Reversing the curse of dimensionality in the visualisation of high-dimensional data

The left hand side is always the grand tour, and the right is the sage tour.

##  Single cell RNA-seq data from mouse retina

Setting sage with $\gamma = 3$

<img src="gifs/mouse_grand.gif" width="450"/> <img src="gifs/mouse_sage_gam3.gif" width="450"/>

## QuickDraw Sketches

<img src="gifs/sketches_grand.gif" width="450"/> <img src="gifs/sketches_sage.gif" width="450"/>

## pdfsense

Setting sage tour with $R = 10$.

<img src="gifs/pdfsense_grand.gif" width="450"/> <img src="gifs/pdfsense_sage_r10.gif" width="450"/>

## pollen: parameter tuning

All sage displays: left is default, middle is $R=1$, and right is $\gamma=20$.

<img src="gifs/pollen_sage.gif" width="300"/> <img src="gifs/pollen_sage_R1.gif" width="300"/> <img src="gifs/pollen_sage_gam20.gif" width="300"/>