### 2020-09-15 paper.Rmd
google bib added year

---
title: |
  Burning sage: Reversing the curse of dimensionality in the visualisation of high-dimensional data
type: Article
author:
  - name: Ursula Laa
    affil: a, b
    email: ursula.laa@monash.edu
  - name: Dianne Cook
    affil: b
    email: dicook@monash.edu
  - name: Stuart Lee
    affil: b, c
    email: stuart.lee1@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
  - num: c
    address: |
      Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: |
  In high-dimensional data analysis the curse of dimensionality reasons that points tend to be far away from the center of the distribution and on the edge of high-dimensional space. Contrary to this, is that projected data tends to clump at the center. This gives a sense that any structure near the center of the projection is obscured, whether this is true or not. A transformation to reverse the curse, is defined in this paper, which uses radial transformations on the projected data. It is integrated seamlessly into the grand tour algorithm, and we have called it a burning sage tour, to indicate that it reverses the curse. The work is implemented into the tourr package in R. Several case studies are included that show how the sage visualisations enhance exploratory clustering and classification problems.
keywords: |
  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(gridExtra)
library(ggforce)
library(tourr)
library(latex2exp)
```

\doublespacing
# Introduction

<!-- Reverse, redistribute, counteract, offset, rescale, radial transformation -->

The term "curse of dimensionality" was originally introduced by @BellmanRichard1961, to express the difficulty of doing optimization in high dimensions because of the exponential growth in space as dimension increases. A way to think about it is that the volume of the space grows exponentially with dimension, which makes it infeasible to sample *enough* points -- any sample will be less densely covering the space as dimension increases. The effect is that most points will be far from the sample mean, on the edge of the sample space. @doi:10.1111/j.1467-9868.2005.00510.x have shown that in the extreme case of high-dimension, low-sample size data, observations are on the vertices of a simplex. This affects many aspects of data analysis: minimizing the error during model fitting relies on effective optimisation techniques, non-parametric modeling requires finding nearest neighbors which may be far away and sampling from high-dimensional distributions is likely to have points far from the population mean.

<!-- This causes problems in different domains. An important example in statistics is cluster analysis. Since the distance between any two points in a distribution becomes approximately constant, new methods are needed to cluster points in a high-dimensional space (see e.g. @Steinbach2004).-->

@Donoho00, conversely, considers the curse of dimensionality as a blessing, because the sparsity can be leveraged for computational efficiency. This is used in regularization methods, like lasso, to penalize model complexity. The penalty results in shrinking (some of) the parameter estimates towards zero.<!-- In ridge regression we obtain a global shrinkage, while Lasso can be thought of as variable selection and will shrink some of the estimates to zero. XXX mention concentration of measures?-->

Paradoxically, the curse of dimensionality inverts for dimension reduction, and affects the resulting visualizations. For low-dimensional linear projections of high-dimensional data, most projections will show approximately Gaussian distributions of the data, with observations concentrating in the center [@diaconis1984]. In projection pursuit, this motivated the development of indexes showing departure from normality. For high-dimension low-sample size data @10.2307/27639976,@10.1093/biomet/asp084 call this "data piling": all observations are in a single point in certain low-dimensional projections. These issues also persist with non-linear dimension reduction techniques. This is often referred to as the "crowding problem", which methods like t-Distributed Stochastic Neighbor Embedding (t-SNE) [@tsne] aim to alleviate.

For illustration we show histograms of two-dimensional linear projections of points sampled uniformly within hyperspheres in $p=3, 10, 100$ dimensions in Figure \ref{fig:density}. All distributions are uniform within the maximum radius $R=1$, but the projected points start piling near the center as $p$ increases.

```{r density, fig.cap="Histogram of two-dimensional projections of 10k points sampled uniformly within a $p$-dimensional hyperspheres, for $p=3, 10, 100$. The fill color shows the logarithm of the bin count, with the highest counts shown in yellow. As $p$ increases the samples concentrate near the center.", out.width="95%", fig.width=6, fig.height=2, fig.align = "center"}

set.seed(12345)

n <- 10000

# sample points, only keep first two components for 2D projection
p3 <- geozoo::sphere.solid.random(3, n)$points[, c(1,2)]
p10 <- geozoo::sphere.solid.random(10, n)$points[, c(1,2)]
p100 <- geozoo::sphere.solid.random(100, n)$points[, c(1,2)]
colnames(p3) <- c("x", "y")
colnames(p10) <- c("x", "y")
colnames(p100) <- c("x", "y")

proj_points <- as_tibble(rbind(p3, p10, p100)) %>%
  mutate(p = factor(c(rep("p = 3", n), rep("p = 10", n), rep("p = 100", n)), levels = c("p = 3", "p = 10", "p = 100")))

ggplot(proj_points, aes(x, y)) +
  geom_hex(bins = 20, aes(fill=log(..count..))) +
  scale_fill_viridis_c() +
  facet_wrap(~p, scales = "free") +
  guides(fill = FALSE) +
  theme_bw() +
  theme(axis.title.x=element_blank(), 
        axis.title.y=element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_blank(),
        aspect.ratio = 1)

```

In this work we address the piling in low-dimensional linear projections, which is in particular a barrier for viewing high-dimensional distributions with tour methods [@As85,@BCAH05]. Tours show interpolated sequences of low-dimensional projections of the data. When exploring a dataset with a tour we can discover features that are only visible from specific viewing angles. However, these features may easily be hidden if most of the observations pile up near the center, with a large fraction of the plotting canvas showing only outlying points. To reverse the effect, we introduce a radial transformation that zooms in near the center of the distribution.

The paper is structured as follows: the radial transformation and its implementation is described in Section \ref{sec:method}, we show applications in Section \ref{sec:application} and conclude in Section \ref{sec:concl}.

<!-- XXX probably remove this, for now keeping it in for reference
see also discussion in https://mc-stan.org/users/documentation/case-studies/curse-dims.html on the average length of multivariate normal vector for large $p$, relation to "concentration of measures".-->

# Burning sage algorithm {#sec:method}

To understand why points tend to be away from the center in the high-dimensional space, but pile near the center in low-dimensional projections, we need to consider the projected volume. It is useful to think of the data as being distributed in a hypersphere, i.e. all data points are within a specified distance from the center. Compared to the common assumption of a distribution in a box or hypercube this has the advantage of being rotation invariant, and more accurately captures commonly observed distributions.

```{r, out.width="60%", fig.cap="The diagram on the left is illustrating the different volumes considered for a three-dimensional sphere. The full sphere has volume $V(R, p)$. Within a radius $r$ the sphere contains the reduced volume $V(r; p, R)$, shown in blue, but the projected volume within a radius $r$ in a two-dimensional plane is much larger, given by the volume of the cylinder with rounded caps, $V_{2D}(r; p ,R)$, shown in red. The intersection of the plane with the sphere is illustrated in grey, and the plane with both radii is shown in the right diagram.", fig.align = "center"}

xp <- 0.4
n <- 200

x <- seq(-xp, xp,length.out = n)
yp <- sqrt(1-x^2)
ym <- -yp

x_outer1 <- c(seq(1-xp, 1, length.out = n/2))
yp_outer1 <- sqrt(1-x_outer1^2)
ym_outer1 <- -yp_outer1

x_outer2 <- c(seq(-1, -1+xp, length.out = n/2))
yp_outer2 <- sqrt(1-x_outer2^2)
ym_outer2 <- -yp_outer2

p1 <- ggplot() +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = 1, b = 0.2, angle = 0), fill = "grey") +
  geom_circle(aes(x0= 0, y0=0, r = 1)) +
  geom_circle(aes(x0= 0, y0=0, r = xp), color="blue", fill = "blue", alpha=0.5) +
  geom_ellipse(aes(x0 = 0, y0 = 0, a = xp, b = xp/10, angle = 0), color="red") +
  geom_ellipse(aes(x0 = 0, y0 = sqrt(1 - (2*xp)^2/4)-xp/100, a = xp, b = xp/10, angle = 0), color="red") +
  geom_ellipse(aes(x0 = 0, y0 = -sqrt(1 - (2*xp)^2/4)+xp/100, a = xp, b = xp/10, angle = 0), color="red") +
  geom_area(aes(x=x, y=yp), fill = "red", alpha = 0.5) +
  geom_area(aes(x=x, y=ym), fill = "red", alpha = 0.5) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 0),
               size = 0.5, arrow = arrow(length = unit(0.1, "inches"))) +
  annotate("text", x = 0.5, y = 0.05, label="R") +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = xp),
               size = 0.5, arrow = arrow(length = unit(0.1, "inches"))) +
  annotate("text", x = -0.05, y = xp/2, label="r") +
  annotate("text", x = 0, y = 0.75, label=TeX("\\textit{V}$_{2D}$\\textit{(r;p,R)}")) + 
  annotate("text", x = 0, y = -xp * 0.75, label=TeX("\\textit{V(r;p,R)}")) + 
  annotate("text", x = 0.65, y = 0.5, label=TeX("\\textit{V(R,p)}")) +
  theme_void() +
  coord_fixed()

p2 <- ggplot() +
  geom_circle(aes(x0= 0, y0=0, r = 1), fill = "grey") +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 0),
               size = 0.5, arrow = arrow(length = unit(0.1, "inches"))) +
  annotate("text", x = 0.5, y = 0.05, label="R") +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = xp),
               size = 0.5, arrow = arrow(length = unit(0.1, "inches"))) +
  annotate("text", x = -0.05, y = xp/2, label="r") +
  geom_circle(aes(x0= 0, y0=0, r = xp), color="red") +
  theme_void() +
  coord_fixed()

grid.arrange(p1, p2, nrow=1)

```

We can picture this for a 3D sphere. In any 2D projection a large fraction of the volume will be projected onto a small region near the center of the resulting disk. This is a consequence of the curved shape, which extends further in the orthogonal direction at the center of the disk, since the projection corresponds to an integration over the orthogonal space. The effect will be further enhanced when projecting onto 1D instead. The same intuition holds as we increase the number of original dimensions that are projected down to one or two dimensions for visualization, with increasing fractions of the volume being projected onto a small area around the center, see Figure \ref{fig:density} for illustration.

To reverse this effect, we introduce a radial transformation that redistributes the projected points, such that equal volume in the original ($p$-dimensional) space is projected onto equal area in a 2-dimensional projection. Note that it is straightforward to generalize this for $d$-dimensional projections by mapping onto equal $d$-dimensional volume instead.

## Definition of the relative projected volume

To understand how the $p$ dimensional volume is projected onto a $2$ dimensional plane, we study what fraction of the total volume is projected onto the area of a disk depending on its radius. This dependence was described in @Laa:2020wkm. We start from a $p$ dimensional hypersphere with radius $R$ and volume $V(R, p)$, and its projected volume onto a centered $2$-dimensional disk of radius $r$, $V_{2D}(r; p, R)$, where $r$ can be any radius within $[0, R]$. The relative projected volume is then given as the ratio of these two quantities,
\begin{equation}
v_{2} (r; p, R) = \frac{V_{2D}(r; p, R)}{V(R, p)} = 1 - \left(1-\left(\frac{r}{R}\right)^2\right)^{p/2}.
\label{eq:cdf}
\end{equation}
This ratio is of particular interest because it gives the 2-dimensional radial cumulative distribution function (CDF) of points when assuming a uniform distribution within the $p$-dimensional hypersphere.

We can compare $v_{2} (r; p, R)$ to the relative volume within a radius $r$ in the original $p$-dimensional hypersphere,
\begin{equation}
v_{p} (r; p, R) = \frac{V(r, p)}{V(R, p)} = \left({\frac{r}{R}}\right)^p.
\end{equation}
This comparison is shown in Fig. \ref{fig:cdf}, for $p=3, 10, 100$. The left graph shows $v_{p} (r; p, R)$ and the right graph $v_{2} (r; p, R)$. We clearly see the opposing behavior of the two functions as $p$ increases. While $v_{p} (r; p, R)$ concentrates more and more towards large values of $r/R$, $v_{2} (r; p, R)$ has the inverse behavior.

```{r cdf, fig.cap="Comparing the relative volume of a $p$ dimensional hypersphere captured within a radius $r$, in the $p$-dimensional space (left) and in a 2-dimensional projection (right), for $p=3, 10, 100$. While most of the volume is squeezed towards larger values of $r$ as $p$ increases, the projected volume gets concentrated near the center instead.", fig.height=2.2, fig.width=6, out.width = "90%", fig.align = "center"}

cdf_2 <- function(p){
  function(r){
    1 - (1 - r^2)^(p/2)
  }
}

cdf_p <- function(p){
  function(r){
    r^p
  }
}

p1 <- ggplot(data = data.frame(r = c(0,1)), mapping = aes(x = r)) +
  stat_function(
        fun = cdf_p(3),
        mapping = aes(color = "ca")) +
  stat_function(
        fun = cdf_p(10),
        mapping = aes(color = "cb")) +
    stat_function(
        fun = cdf_p(100),
        mapping = aes(color = "cc")) +
  scale_color_manual(name = "p",
                     values = RColorBrewer::brewer.pal(3, "Dark2"),
                     labels = c("3", "10", "100")) +
  xlab("r/R") + ylab("Relative volume") + theme_bw() +
  ggtitle("Volume in p dimensions") +
  theme(legend.position=c(0.2, 0.55))

p2 <- ggplot(data = data.frame(r = c(0,1)), mapping = aes(x = r)) +
  stat_function(
        fun = cdf_2(3),
        mapping = aes(color = "ca")) +
  stat_function(
        fun = cdf_2(10),
        mapping = aes(color = "cb")) +
    stat_function(
        fun = cdf_2(100),
        mapping = aes(color = "cc")) +
  scale_color_manual(name = "p",
                     values = RColorBrewer::brewer.pal(3, "Dark2"),
                     labels = c("3", "10", "100")) +
  xlab("r/R") + ylab("Relative volume") + theme_bw() +
  ggtitle("2D projected volume") +
  guides(color = FALSE)

grid.arrange(p1, p2, ncol=2)

```

## Calculating the radial transformation

The aim of the algorithm is to redistribute the projected volume such that equal relative areas on the disk, as given by $v_{2} (r; p=2, R)= v_p(r; p=2, R) = (r/R)^2$, contain equal relative projected volume, given by $v_{2} (r; p, R)$. This is achieved through a transformation of the projected radius that can be defined for any $r\in[0,R]$, and is applied to the projected data points in the plane, $y = (y_1, y_2)$. We work with polar coordinates and represent the data points as $y = (r_y, \theta_y)$. The angular component $\theta_y$ is uniform for this distribution, by the rotation invariance of the sphere, and thus does not need to be transformed. The radial component $r_y$ is transformed in two steps.

The first transformation is to replace $r_y$ with $v_{2} (r_y; p, R)$. Since this is the radial CDF of the assumed underlying distribution, we expect that $v_{2} (r_y; p, R)$ is approximately uniformly distributed in radius. We then transform $v_{2} (r_y; p, R)$ using the inverse of $v_{2} (r_y; 2, R)$, to go from a uniform distribution in radius to a uniform distribution in area of the disk. This inverse is defined via
\begin{equation}
v_2^{-1}(v_2(r_y; 2, R); 2, R) = v_2(v_2^{-1}(r_y; 2, R); 2, R) = r_y
\end{equation}
and thus
\begin{equation}
v_2^{-1}(r_y; 2, R) = R \sqrt{r_y}.
\end{equation}

The full radial transformation is therefore given by 
\begin{equation}
r'_y = v_2^{-1} (v_2(r_y; p, R); 2, R) =  R \sqrt{v_2(r_y; p, R)} = R \sqrt{1-\left(1-\left(\frac{r_y}{R}\right)^2\right)^{p/2}}.
\label{eq:resc}
\end{equation}

The relation between $r'_y$ and $r_y$ depends on the number of dimensions $p$, and is illustrated for selected values in Figure \ref{fig:radii}. We see that the transformation is approximately linear near the center. As $p$ increases it becomes non-linear faster, and e.g. for $p=10$ the points with radius $r_y>0.5$ will already be highly distorted and pushed out towards the last eighth in $r'_y$. Figure \ref{fig:circles} demonstrates this for different values of $p$ by showing equidistant circles for which the radius has been transformed according to Eq. \ref{eq:resc}.

```{r radii, fig.cap="Relation between $r_y$ and $r'_y$ for different values of $p$ and assuming $R=1$. The scaling is approximately linear near the center, but leads to distortion at large radii when $p$ is large.", fig.height=3, fig.width=6, out.width="60%", fig.align = "center"}
# define index as function of c
trans_p <- function(p){
  function(r){
    sqrt( 1 - (1-r^2)^(p/2))
  }
}

# plot dependence for selected values
ggplot(data = data.frame(x = c(0,1)), mapping = aes(x = x)) +
    stat_function(
        fun = trans_p(2),
        mapping = aes(color = "ca")) +
  stat_function(
        fun = trans_p(4),
        mapping = aes(color = "cb")) +
  stat_function(
        fun = trans_p(6),
        mapping = aes(color = "cc")) +
    stat_function(
        fun = trans_p(10),
        mapping = aes(color = "cd")) +
    stat_function(
        fun = trans_p(20),
        mapping = aes(color = "ce")) +
  scale_color_manual(name = "p",
                     values = RColorBrewer::brewer.pal(5, "Dark2"),
                     labels = c("2 (baseline)", "4", "6", "10", "20")) +
  xlab(expression(r[y])) + ylab(expression(r[y]*minute)) + theme_bw()

```

```{r circles, fig.cap="Equidistant concentric circles, for $p=2, 3, 10, 100$, illustrating the radial transformation. The circles remain equidistant for $p=2$ where no transformation is performed, and get pushed out towards the edge as $p$ increases.", out.width="75%", fig.align = "center"}

lbl <- c("p=2", "p=3", "p=10", "p=100")
names(lbl) <- c("p4", "p6", "p10", "p20")

tibble(x = 0, y = 0, r = seq(0.1, 1, length.out = 10)) %>%
  mutate(p4 = trans_p(2)(r)) %>%
  mutate(p6 = trans_p(3)(r)) %>%
  mutate(p10 = trans_p(10)(r)) %>%
  mutate(p20 = trans_p(100)(r)) %>%
  mutate(r = as.factor(r)) %>%
  pivot_longer(cols = starts_with(("p"))) %>%
  mutate(name = factor(name, levels = c("p4", "p6", "p10", "p20"))) %>%
  ggplot() +
  geom_circle(aes(x0=x, y0=y, r=value, color=r)) +
  coord_fixed() +
  theme_bw() +
  facet_wrap(~name, labeller = labeller(name = lbl)) +
  guides(color = FALSE) +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())

```

## Trimming and tuning {#sec:params}

The transformation in Eq. \ref{eq:resc} is fixed for a given input dataset, by evaluating the number of dimensions $p$ and the maximum distance from the center $R$. However, in realistic applications we may wish to trim the projected data or tune the transformation. In practice, a combination of both adjustments can be used to further zoom in on the center of the distribution, or alternatively, to soften the transformation.

### Trimming
The overall scale of the transformation is determined by $R$. In the case of an approximately spherical and uniform distribution the maximum distance from the center works well and ensures the validity of the rescaling in Eq. \ref{eq:resc}. But this is not robust and might result in a much larger scale than desired, especially when it is determined by outlying observations.

We therefore allow trimming of the projected observations, using $R$ as a free parameter of the display function. When selecting a value $R$ that is smaller than the maximum distance from the center, we need to ensure that the projected radius of points is always smaller than $R$, by trimming $r_y$ as
\begin{equation}
r_y^{\mathrm{trim}} = \min(r_y, R)
\label{eq:cutR}
\end{equation}
for each observation.

### Tuning
The dimension of the input might not reflect the intrinsic dimensionality of the dataset. This could be the case when dimension reduction was used prior to visualization, e.g. displaying only the first few principal components. In this case the effective dimensionality $p_{\mathrm{eff}}$ is likely between the original number of dimensions and the selected number of principal components. We can think of omitted components as being in the orthogonal space of all considered projections, with some directions being pure noise, while others may still carry relevant information.

We allow tuning $p_{\mathrm{eff}} = \gamma p$ by selecting the scaling parameter $\gamma$. By default, $\gamma=1$ and $p_{\mathrm{eff}}=p$. When $\gamma<1$ the rescaling will be softer, and $\gamma>1$ results in more aggressive rescaling than suggested by $p$ alone. Note that when $p_{\mathrm{eff}} < 2$ we actually invert the behavior and shift the focus away from the center, in general this is not wanted.

## Implementation as a dynamic display {#sec:implementation}

While the radial transformation can in general be used with any low-dimensional display that suffers from data piling, it is most useful when combined with a dynamic display showing a sequence of interpolated low-dimensional projections obtained when running a tour. We have implemented it as a new display method `display_sage` in the `tourr` package [@tourr] in R [@rref].

We can think of the display functions as part of a data pipeline obtained when running a tour. The initial step is pre-processing the data, given by $\mathbf{X}$, an $n \times p$ matrix containing $n$ observations in $p$ dimensions. Typically, this includes centering and scaling, using either the overall range or the variance. Ensuring a common scale of all variables, comparable to the selected scale parameter $R$, is especially important with the new display. The tour is then looping over the following steps:

1. Obtain projection matrix $\mathbf{A}$. For $d$-dimensional projections this is an orthonormal $p \times d$ matrix. To ensure the smooth rotation of projections, each new $\mathbf{A}$ is obtained as an interpolated step in the sequence, as explained in @BCAH05.
2. Project the data by computing $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$.
3. Map $\mathbf{Y}$ to the display to re-draw the projected data. For $d=2$ this typically maps the projected points onto a scatter plot display. With the new display we first transform $\mathbf{Y}$ as:
    + Center the 2-dimensional matrix $\mathbf{Y}$ and compute its polar coordinate representation $(r_y, \theta_y)$.
    + For each observation, first use Eq. \ref{eq:cutR} to get the trimmed radius $r_y^{\mathrm{trim}}$  within the specified range, and then apply the radial transformation defined in Eq. \ref{eq:resc} to obtain $r'_y$.
    + Use the transformed radial coordinate $r'_y$ to re-compute the mapping onto Euclidean coordinates $(y_1, y_2)$.
4. To fit the final projection onto the plotting canvas ranging between $[-1,1]$, we rescale each mapped observation $y$ using a scaling parameter $s$.

The display can be added when calling the `animate` function in `tourr`, as

```{r codeExample, echo=TRUE, eval=FALSE}
tourr::animate(
  data,
  tour_path = tourr::grand_tour(),
  display = display_sage(gam, R, half_range)
  )
```

and uses `gam` to set the $\gamma$ parameter for computing $p_{\mathrm{eff}}$, and the overall range `R` used for trimming. Both these parameters are described in Section \ref{sec:params} above. Finally, `half_range` sets the scale parameter $s$ to adjust the scale to the drawing canvas. The ratio $R/s$ sets the scale for fitting the displayed data on the plotting canvas, by default $s = R$ and we apply the scaling factor $0.9$ to contain the projected points within the display. When adjusting $R$ the user should take care to adjust $s$ accordingly.

# Applications {#sec:application}

Four applications of the sage tour are shown here, which illustrate the benefit of using the reverse transformation for examining data using a tour: single cell RNA-seq, classifying hand-sketched images, comparing physics experiments, and the classical pollen data which is used to illustrate the effect of parameter choices.

## Clustering single-cell RNA-seq data sets

In the analysis of single cell RNA-seq data, cluster analysis is performed to
detect cell types and find expression of genes that define those cell types, and the relative orientation of the cell types  to each other (trajectory analysis). Generally, for cluster verification, analysts use embedding methods
like t-SNE to verify the placement and meaning of clusters from a clustering algorithm.

Here we compare the sage display and regular tour display on mouse retinal single cell RNA-seq data from
@Macosko2015-ot. The raw data consists of a 49,300  cells and was downloaded
using the using the **scRNAseq** Bioconductor package [@scRNAseq-d]. We
use a standard workflow for pre-processing and normalizing this data 
(described by @Amezquita2020-at): we performed QC using the **scater** package
and we log-transformed and normalised the expression 
values and selected highly variable genes (HVGs) using **scran** 
[@McCarthy2017; @Lun2016]. The top ten percent of HVGs were used to subset
the normalised expression matrix and compute PCA using the first 25 components.
Using the PCs we built a shared nearest neighbours graph (with $k = 10$) and 
used Louvain clustering which generates 11 clusters [@Blondel2008-bx].

We tour the first five PCs (approximately 20% of the variance in expression),
on a weighted subsample of cells based on their cluster membership - 4,590 cells. When running the sage display, we set $\gamma = 3$ as the effective dimensionality of the data. For the displays we have rescaled the PCs so they each have zero mean and unit variance. Here we are interested in mitigating the effects of overplotting by focusing on three clusters that were identified by graph based clustering but appear to overlap on PC plots.

```{r mouse-cluster, fig.cap = "Selected frames from using a tour of the mouse data with the default tour display (top), and the sage display with $\\gamma=3$ (bottom)", out.width="0.9\\textwidth", fig.align="center"}
library(magick)
dir <- "pngs/"
frames <- c("001", "016", "038", "074")
draw_frames <- function(pngs) {
  res <- lapply(pngs, image_read)
  res <- do.call("c", res)
  image_append(res)
}

draw_panel <- function(a, b) {
  whitespace <- image_blank(image_info(a)$width, 50, color = "none")
  img <- c(a, whitespace, b)
  image_append(img, stack = TRUE)
}
mouse_grand <- paste0(dir, "mouse_grand_2c-", frames, ".png")
mouse_grand_panel <- draw_frames(mouse_grand)
mouse_sage <- paste0(dir, "mouse_sage_2c_gam3-", frames, ".png")
mouse_sage_panel <- draw_frames(mouse_sage)

mouse <- draw_panel(mouse_grand_panel, mouse_sage_panel)
mouse
```

Using the default tour display (figure \ref{fig:mouse-cluster}, top), the three clusters (dark green, blue, and yellow) are obscured by points in other clusters as we move through the frames of the animation. The points in dark green cluster are overlapping those found in the yellow and blue clusters; and it is difficult to see if there is any separation between the blue and yellow clusters.

On the other hand using the sage display (figure \ref{fig:mouse-cluster}, bottom), expands the center of projection, and results in the
three clusters appearing more separable than the default display. The dark green cluster becomes more visible; the outlier cell that is far away from the rest of the clustered points is more apparent. Moreover, the relative positions of the yellow and blue clusters centers are easier to see, those clusters are distinct from the dark green cluster, yet in most frames they appear to overlap and jumble together, indicating that perhaps those clusters should be merged together.

## Classifying hand-sketches

We can use the new display to look at different distributions of images from Google QuickDraw collection [@quickdraw-paper]. These are $28\times28=784$ pixel grey scale data that are available publicly. In this example, we sample 1000 images from three types of sketches (banana, cactus, crab) and see if we can separate the classes in the high-dimensional parameter space.

We reduce the dimensionality from 784 variables to the first 5 PCs, which captures approximately 20 per cent of the variation of the data. Before applying the tour we rescale each component to have mean zero and unit variance.

```{r sketches, fig.cap = "Selected frames  of the tour run on the sketches data using the default tour display (top), and using the default sage display (bottom).", out.width="0.9\\textwidth", fig.align="center"}
frames <- c("001", "036", "077", "092")

sketches_grand <- paste0(dir, "sketches_grand-", frames, ".png")
sketches_grand_panel <- draw_frames(sketches_grand)
sketches_sage <- paste0(dir, "sketches_sage-", frames, ".png")
sketches_sage_panel <- draw_frames(sketches_sage)

sketches <- draw_panel(sketches_grand_panel, sketches_sage_panel)
sketches
```

Figure \ref{fig:sketches} shows the grand tour on the PCs, where green points correspond to the banana class, orange points represent the cactus class and purple points are the crab class. In both displays points belonging to the cactus class are concentrated in the center, however on the default display (figure \ref{fig:sketches}, top) there is overplotting: points from other classes overlap those in the cactus class. The sage display (figure \ref{fig:sketches}, bottom) helps reduce overplotting, it is easier to see that the centers of class are separated and that there is substructure in the banana class, which further collapses into two subgroups.

## Comparing physics experiments: PDFSense

Data were obtained from CT14HERA2 parton distribution function fits and describe the sensitivity of fit parameters to experimental measurements [@Wang:2018heo]. There are 28 parameters, and varying one at a time to move $\pm 1 \sigma$ away from the 'best fit point' (maximum likelihood estimate) provides our input variables, labelled X1-X56. Each observation corresponds to the fit prediction for these 56 parameter combination for a single physical observable. Points are grouped based on the underlying process in the experiment, which is mapped to color in the following. With the analysis of the distribution along these variables X1-X56 we can understand to what extent each experimental measurement provides complementary information for the global fit, e.g. orthogonality between the three groups or the identification of outlying points which are considered as important for future fits, see discussion in @Cook:2018mvr.

```{r pdfsense, fig.cap = "Selected frames of the tour run on the pdfsense data using the default tour display (top), and using the sage display with $R=10$ (bottom)", fig.align="center", out.width="0.9\\textwidth"}
frames <- c("001", "051", "078", "092")

pdfsense_grand <- paste0(dir, "pdfsense_grand-", frames, ".png")
pdfsense_grand_panel <- draw_frames(pdfsense_grand)
pdfsense_sage <- paste0(dir, "pdfsense_sage_r10-", frames, ".png")
pdfsense_sage_panel <- draw_frames(pdfsense_sage)

pdfsense <- draw_panel(pdfsense_grand_panel, pdfsense_sage_panel)
pdfsense
```

Following the processing described there, we tour the first 6 principal components, rescaled to have zero mean and unit variance. In figure \ref{fig:pdfsense} we see that using the default tour display (figure \ref{fig:pdfsense}, top) and the sage display with $R = 10$ (figure \ref{fig:pdfsense}, bottom), maintains the overall shape of the data (different physical process are orthogonal in the parameter space). The particular structure of this distribution, with some clusters extending linearly away from the center and a set of outlying points, results in poor use of the plotting space, and high level of clustering near the center. For example, focussing on the blue cluster, we can see that it extends out along different directions, but it can be challenging to observe how the point move under the tour rotation, as overplotting becomes an issue when points move through the center. Here, the new display has a clear advantage, as illustrated with the selected views in figure \ref{fig:pdfsense} (bottom).

## Tuning the parameters: Pollen

The classical pollen data allows us to better demonstrate the use of the trimming and tuning parameters. The five dimensional data set was simulated by David Coleman of RCA Labs, for the Joint Statistics Meetings 1986 Data Expo [@pollen], and is an example of a hidden structure near the centre of a distribution. The data are standardised by centering and scaling such that the standard deviation of each variable is equal to one.

Neither the standard scatter display nor the sage display with default settings can reveal the structure, see still view from the sage display with default settings in Figure \ref{fig:pollen} (left,  using the maximum radius in the scaled data, thus $(R=6.6, \gamma=1)$). We next use either $\gamma$, $R$ or a combination of the two to zoom in further near the center, see the middle $(R=1, \gamma=1)$ and right $(R=6.6,\gamma=20)$ plots in Figure \ref{fig:pollen} for examples. There is approximate equivalence between the results obtained using either tuning or trimming, and both views clearly reveal the word "EUREKA" hidden in the distribution.

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{png/pollen_sage-88.png}
\includegraphics[width=0.3\textwidth]{png/pollen_sage_R1-88.png}
\includegraphics[width=0.3\textwidth]{png/pollen_sage_gam20-88.png}
\caption{Selected view of the pollen data in the new sage display, with default settings (left), setting $R=1$ (middle) or $\gamma=20$ (right). We can tune either $\gamma$, $R$ or a combination of the two to reveal the word "EUREKA" near the center of the distribution.}
\label{fig:pollen}
\end{figure}

# Discussion {#sec:concl}

In this paper we have introduced the sage display, as a visualisation technique for reversing
the curse of dimensionality in high dimensional projections, and have shown its utility in applications to analysis tasks found in the natural sciences and machine learning. Our case studies highlight that sage display can reduce the effects of overplotting - in the sketches example our new display decreases the number of overlapping points between classes and provides
better visual separation between classes compared to the regular tour. Furthermore, unlike some
non-linear dimension reduction (NLDR) techniques like t-SNE, the sage display still gives the viewer an accurate assessment of the size of a cluster, as we saw in our single cell example. Our parameter tuning example highlights that the sage display is linear up to a given effective dimension. As a result, the sage display provides interpretable visualisations while preserving global structure in the high dimensional space, as highlighted in the particle physics example.

The recently introduced slice tour [@sliceTour] is essentially addressing the same issue, i.e. the overplotting of points from the large orthogonal space, but follows a different philosophy. Instead of mapping orthogonal volume to projected volume, we cut through the orthogonal space and only select a small portion of the volume for plotting. This is done by defining a slice via a cutoff on the orthogonal distance of points. The approach would be preferred in the case of large number of points, or for concave structure in the distribution, while the sage display works better in the case of limited sample size.

For the sage display tuning is often required to avoid bunching points in the visualisation, especially if the projection is dense in the center of the distribution. However, given that the display is quick to compute, it is easy to try out different parameter combinations. Expanding around the center too aggressively can also introduce distortions in the display, which is also a problem for NLDR methods.

Currently, the sage display is naive to information in high dimensional space that is not captured by linear projections. Further work is required to investigate other transformations of the projections that can capture non-linear structure or that are locally linear, like the transformations performed by t-SNE.

<!-- Add note on interactivity to tuning paragraph. Add a paragraph on applications -->