### 2020-09-16 supp-mouse.Rmd
reformat supplementary material, add materials as zip file

---
title: "Supplementary Materials - Single Cell Mouse Retina Data"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  error = FALSE)
```

##  Single cell RNA-seq data from mouse retina

The left hand side is the regular tour
display, and the right hand side is sage display with $\gamma = 3$.

<img src="gifs/mouse_grand_2c.gif" width="450"/> <img src="gifs/mouse_sage_2c_gam3.gif" width="450"/>