### 2020-09-16 supp-pollen.Rmd

---
title: "Supplementary Materials - Needle in a haystack, the pollen dataset"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  error = FALSE)
```

## pollen: parameter tuning

All sage displays: left is default, middle is $R=1$, and right is $\gamma=20$.

<img src="gifs/pollen_sage.gif" width="300"/> <img src="gifs/pollen_sage_R1.gif" width="300"/> <img src="gifs/pollen_sage_gam20.gif" width="300"/>