### 2021-07-15 paper/paper.tex
Working on the revision

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% see https://stackoverflow.com/a/47122900
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% Pandoc citation processing

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}

\begin{document}

\articletype{Article}

\title{Burning sage: Reversing the curse of dimensionality in the
visualization of high-dimensional data}

\author{\name{Ursula Laa$^{a, b}$, Dianne Cook$^{b}$, Stuart Lee$^{b,
c}$}
\affil{$^{a}$School of Physics and Astronomy, Monash
University; $^{b}$Department of Econometrics and Business Statistics,
Monash University; $^{c}$Molecular Medicine Division, Walter and Eliza
Hall Institute, Parkville, Australia}
}

\thanks{CONTACT Ursula
Laa. Email: \href{mailto:ursula.laa@monash.edu}{\nolinkurl{ursula.laa@monash.edu}}, Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}, Stuart
Lee. Email: \href{mailto:stuart.lee1@monash.edu}{\nolinkurl{stuart.lee1@monash.edu}}}

\maketitle

\begin{abstract}
In high-dimensional data analysis the curse of dimensionality reasons
that points tend to be far away from the center of the distribution and
on the edge of high-dimensional space. Contrary to this, is that
projected data tends to clump at the center. This gives a sense that any
structure near the center of the projection is obscured, whether this is
true or not. A geometric transformation to reverse the curse, is defined
in this paper, which uses radial transformations on the projected data.
It is integrated seamlessly into the grand tour algorithm, and we have
called it a burning sage tour, to indicate that it reverses the curse.
The work is implemented into the tourr package in R. Several case
studies are included that show how the sage visualizations enhance
exploratory clustering and classification problems.
\end{abstract}

\begin{keywords}
data visualisation; grand tour; statistical computing; statistical
graphics; multivariate data; dynamic graphics; data science; machine
learning
\end{keywords}

\doublespacing

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The term ``curse of dimensionality'' was originally introduced by
\citet{BellmanRichard1961}, to express the difficulty of doing
optimization in high dimensions because of the exponential growth in
space as dimension increases. A way to think about it is, that the
volume of the space grows exponentially with dimension, which makes it
infeasible to sample \emph{enough} points -- any sample will be less
densely covering the space as dimension increases. The effect is that
most points will be far from the sample mean, on the edge of the sample
space. \citet{doi:10.1111/j.1467-9868.2005.00510.x} have shown that in
the extreme case of high-dimension, low-sample size data, observations
are on the vertices of a simplex.

This affects many aspects of data analysis: minimizing the error during
model fitting relies on effective optimization techniques,
non-parametric modeling requires finding nearest neighbors which may be
far away and sampling from high-dimensional distributions is likely to
have points far from the population mean. \citet{Donoho00} considers the
curse of dimensionality as a blessing, because the sparsity can be
leveraged for computational efficiency. This is used in regularization
methods, like lasso, to penalize model complexity. The penalty term
results in shrinking (some of) the parameter estimates towards zero.

Paradoxically, the curse of dimensionality inverts for dimension
reduction, resulting in an excessive amount of observations near the
center of the distribution. This affects visualizations made on
low-dimensional projections. The effect is described by
\citet{diaconis1984},
\textcolor{red}{showing that most linear projections onto low-dimensional subspaces are approximately Gaussian, and that observations concentrate near the center as the dimensionality of the full space increases.}
This has motivated the development of indexes for projection pursuit
which search for departure from normality. It is also related to what is
called ``data piling'' in high-dimension low-sample size data
\citep{10.2307/27639976, 10.1093/biomet/asp084}: all observations can
collapse into a single point. These issues also persist with non-linear
dimension reduction techniques, and are often referred to as the
``crowding problem'', which methods like t-Distributed Stochastic
Neighbor Embedding (t-SNE) \citep{tsne} aim to alleviate.

Here we are providing a solution to fix a long-standing problem in the
tour \citep{As85, BCAH05}, where high-dimensional data is viewed by a
movie of low-dimensional projections. Points crowding the center of the
drawing canvas detract from reading the plots, even with relatively few
dimensions. This is a structural problem related purely to the geometry
of projections from high-dimensional spaces. We show that it can be
alleviated with a radial transformation on the low-dimensional space.
Figure \ref{fig:density} illustrates this crowding problem, as would be
experienced when using a tour. Two-dimensional linear projections of
data, generated by sampling uniformly within \(p\)-dimensional
hyperspheres (\(p=3, 10, 100\)), are displayed as hexbin plots. Color
indicates log count of the bin, with yellow mapping the highest counts.
As \(p\) increases the density concentrates in the center of the
projection. (Note that, sampling from a uniform in a sphere is used for
this illustration for several reasons. Any linear projection is
equivalent because of the rotation invariance of the distribution. The
uniform sampling places focus on the space, points are spread evenly
across the volume, and not a data distribution. Also, \(p=10\) would
already be considered high-dimensional, as seen by the piling observed
in the middle plot.)

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{paper_files/figure-latex/density-1}

}

\caption{Illustration of data crowding, using hexbin plots of two-dimensional projections of 10k points sampled uniformly within $p$-dimensional hyperspheres, for $p=3, 10, 100$. The fill color shows the logarithm of the bin count, with the highest counts shown in yellow. As $p$ increases the density  concentrates near the center.}\label{fig:density}
\end{figure}

Common solutions to overplotting include transparency and binned or
density displays \citep{chang}. These are designed to tackle large
sample size, rather than dimension reduction problems. A low-dimensional
projection represented as a scatterplot, binned plot, or contour plot,
will still suffer from data piling and reduced resolution near the
center of the projection. These approaches are still useful for large
sample sizes, and could be applied after the transformation.

The underlying idea of the (radial) transformation, to reverse the
curse, is to correct for the distortion in geometry that occurs when
projecting from a high-dimensional space onto low dimensions. This is
done by mapping equal volume in the full space onto equal area in a
two-dimensional projection via a geometrical transformation of the
projected data. When used with data that is uniformly distributed in the
high-dimensional space, the transformation will result in a uniform
distribution in the low-dimensional display. In general, for
\textit{any} underlying data distribution, we can instead think about
the effect in terms of resolution: adjusting for the geometric piling
effect means that the resolution depends on the projected volume. The
observed projected distribution is representative of the original
high-dimensional distribution in this sense.

The paper is structured as follows. The radial transformation and its
implementation is described in Section \ref{sec:method}. Section
\ref{sec:application} illustrates the use of the sage tour with examples
in clustering, supervised classification and a classical
needle-in-the-haystack problem. Section \ref{sec:concl} describes
possible extensions to the method.

\hypertarget{sec:method}{%
\section{Burning sage algorithm}\label{sec:method}}

To understand why points tend to be away from the center in the
high-dimensional space, but crowd the center in low-dimensional
projections, it is helpful to consider the projected volume relative to
high-dimensional volume. To avoid edge effects and to impose rotation
invariance, we start from the volume in a hypersphere, i.e.~the volume
is contained within a specified distance from the center. This is more
tractable than assuming a hypercube (box).

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{paper_files/figure-latex/sketch-1}

}

\caption{Illustration (and notation) for describing the elements used in the burning sage transformation. The 3D sphere (left) shows the different volumes to be compared. The full sphere has volume $V(R, p)$. Within a radius $r$ the sphere contains the reduced volume $V(r; p, R)$, shown in blue, but the projected volume within a radius $r$ in a two-dimensional plane is much larger, given by the volume of the cylinder with rounded caps, $V_{2D}(r; p ,R)$, shown in red. The intersection of the plane with the sphere is illustrated in grey, and the plane representing the projection with both radii is shown at right.}\label{fig:sketch}
\end{figure}

Figure \ref{fig:sketch} illustrates the comparison to be made, using
something we can easily picture, a 3D sphere. Projecting the data from
within a 3D sphere to 2D (grey disk) will result in mass being condensed
into the disk. Imagine comparing the volume of a cylinder at different
locations in the disk. A centered cylinder has more volume. This is
exaggerated as \(p\) increases: the centered cylinder has much more
volume than any other cylinder.

To reverse this effect, we introduce a radial transformation that
redistributes the projected points, such that equal volume in the
original (\(p\)-dimensional) space is projected onto equal area in a
2-dimensional projection. Note that this can be generalized for
\(d\)-dimensional projections by mapping onto equal \(d\)-dimensional
volume instead.

\hypertarget{definition-of-the-relative-projected-volume}{%
\subsection{Definition of the relative projected
volume}\label{definition-of-the-relative-projected-volume}}

To understand how the \(p\)-dimensional volume is projected onto a
\(2\)-dimensional plane, we study what fraction of the total volume is
projected onto the area of a disk depending on its radius. This
dependence was described in \citet{Laa:2020wkm}. We start from a
\(p\)-dimensional hypersphere, with radius \(R\) and volume \(V(R, p)\),
and its projected volume onto a centered \(2\)-dimensional disk of
radius \(r\), \(V_{2D}(r; p, R)\), where \(r\) can be any radius within
\([0, R]\). The relative projected volume is then given as the ratio of
these two quantities, \begin{equation}
v_{2} (r; p, R) = \frac{V_{2D}(r; p, R)}{V(R, p)} = 1 - \left(1-\left(\frac{r}{R}\right)^2\right)^{p/2}.
\label{eq:cdf}
\end{equation}

We can compare \(v_{2} (r; p, R)\) to the relative volume within a
radius \(r\) in the original \(p\)-dimensional hypersphere,
\begin{equation}
v_{p} (r; p, R) = \frac{V(r, p)}{V(R, p)} = \left({\frac{r}{R}}\right)^p.
\label{eq:vp}
\end{equation}

Figure \ref{fig:cdf} compares these two quantities (Eq. \ref{eq:cdf} and
\ref{eq:vp}), for \(p=3, 10, 100\). On the left is \(v_{2} (r; p, R)\)
and on the right is \(v_{p} (r; p, R)\). The function shapes change in
opposite directions as \(p\) increases: \(v_{2} (r; p, R)\) peaks
earlier and \(v_{p} (r; p, R)\) gets flatter. This is the paradox of the
curse of dimensionality in that projected volume at the center increases
with \(p\).

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/cdf-1}

}

\caption{Comparing relative volume of a $p$-dimensional hypersphere captured within a radius $r$, in the 2-dimensional projection (left) and in the $p$-dimensional space (right), for $p=3, 10, 100$. The difference is dramatic, which illustrates the paradox of the curse of dimensionality. The relative volume of a $p$D sphere shrinks, as $p$ increases, while the projected volume (near the center) grows.}\label{fig:cdf}
\end{figure}

\hypertarget{calculating-the-radial-transformation}{%
\subsection{Calculating the radial
transformation}\label{calculating-the-radial-transformation}}

The aim of the algorithm is to redistribute the projected volume such
that equal relative areas on the disk, as given by
\(v_{2} (r; p=2, R)= v_p(r; p=2, R) = (r/R)^2\), contain equal relative
projected volume, given by \(v_{2} (r; p, R)\). This is achieved through
a transformation of the projected radius that can be defined for any
\(r\in[0,R]\), and is applied to the projected data points in the plane,
\(y = (y_1, y_2)\). We work with polar coordinates and represent the
data points as \(y = (r_y, \theta_y)\). The angular component
\(\theta_y\) is uniform for this distribution, by the rotation
invariance of the sphere, and thus does not need to be transformed. The
radial component \(r_y\) is transformed in two steps.

\textcolor{red}{We transform $r_y$ in two steps. First, we replace it with $v_{2} (r_y; p, R)$ to redistribute the volume uniformly by the radius. The second step is then to }
transform \(v_{2} (r_y; p, R)\) using the inverse of
\(v_{2} (r_y; 2, R)\), to go from a uniform distribution in radius to a
uniform distribution in area of the disk. This inverse is defined via
\(v_2^{-1}(v_2(r_y; 2, R); 2, R) = v_2(v_2^{-1}(r_y; 2, R); 2, R) = r_y\)
and thus \(v_2^{-1}(r_y; 2, R) = R \sqrt{r_y}\).

The full radial transformation is therefore given by \begin{equation}
r'_y = v_2^{-1} (v_2(r_y; p, R); 2, R) =  R \sqrt{v_2(r_y; p, R)} = R \sqrt{1-\left(1-\left(\frac{r_y}{R}\right)^2\right)^{p/2}}
\label{eq:resc}
\end{equation}

The relation between \(r'_y\) and \(r_y\) depends on the number of
dimensions \(p\), and is illustrated for selected values in Figure
\ref{fig:radii}. We see that the transformation is approximately linear
near the center. As \(p\) increases it becomes non-linear faster, and
for \(p=10\), for example, the points with radius \(r_y>0.5\) will
already be highly distorted and pushed out towards the last eighth in
\(r'_y\). Figure \ref{fig:circles} demonstrates this for different
values of \(p\) by showing equidistant circles for which the radius has
been transformed according to Eq. \ref{eq:resc}.

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{paper_files/figure-latex/radii-1}

}

\caption{Relation between $r_y$ and $r'_y$ for different values of $p$ and assuming $R=1$. The scaling is approximately linear near the center, but leads to distortion at large radii when $p$ is large.}\label{fig:radii}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/circles-1}

}

\caption{Concentric circles that are equidistant before the application of the radial transformation, for $p=2, 3, 10, 100$, illustrating the effect of the transformation with increasing number of dimensions. The circles remain equidistant for $p=2$ where no transformation is performed, and get pushed out towards the edge as $p$ increases.}\label{fig:circles}
\end{figure}

\textcolor{red}{At this point a few characteristics of the transformation should be noted. First, as already mentioned, it is approximately linear near the center, and thus will not affect shapes in that part of the graph. In addition, since only the radial component is transformed, angles between the points will not be altered. (This can also be shown explicitly, from mathematical reformulation of the angles between two points after the transformation.) Note however that shapes will be affected further away from the center (in particular towards the maximum range $R$). For example, a linear relation might appear curved because of the non-linear nature of the transformation in that region of the projected space. Thus, while the interpretation is straightforward for points projected in the central region, care must be taken with the interpretation of features observed in the outer region of the projected distribution.}

\hypertarget{sec:params}{%
\subsection{Trimming and tuning}\label{sec:params}}

The transformation in Eq. \ref{eq:resc} is fixed for a given input
dataset, by evaluating the number of dimensions \(p\) and the maximum
distance from the center \(R\). However, in practice we may wish to trim
the projected data or tune the transformation. A combination of both
adjustments can be used to further zoom in on the center of the
distribution, or alternatively, to soften the transformation.

\hypertarget{trimming}{%
\subsubsection{Trimming}\label{trimming}}

The overall scale of the transformation is determined by \(R\). In the
case of an approximately spherical and uniform distribution the maximum
distance from the center works well and ensures the validity of the
rescaling in Eq. \ref{eq:resc}. But this is not robust and might result
in a much larger scale than desired, especially when it is determined by
outlying observations.

We therefore allow trimming of the projected observations, using \(R\)
as a free parameter of the display function. When selecting a value
\(R\) that is smaller than the maximum distance from the center, we need
to ensure that the projected radius of points is always smaller than
\(R\), by trimming \(r_y\) as \begin{equation}
r_y^{\mathrm{trim}} = \min(r_y, R)
\label{eq:cutR}
\end{equation} for each observation.

\hypertarget{tuning}{%
\subsubsection{Tuning}\label{tuning}}

The dimension of the input might not reflect the intrinsic
dimensionality of the dataset. This could be the case when dimension
reduction was used prior to visualization, e.g.~displaying only the
first few principal components. In this case the effective
dimensionality \(p_{\mathrm{eff}}\) is likely between the original
number of dimensions and the selected number of principal components. We
can think of omitted components as being in the orthogonal space of all
considered projections, with some directions being pure noise, while
others may still carry relevant information.

We allow tuning \(p_{\mathrm{eff}} = \gamma p\) by selecting the scaling
parameter \(\gamma\). By default, \(\gamma=1\) and
\(p_{\mathrm{eff}}=p\). When \(\gamma<1\) the rescaling will be softer,
and \(\gamma>1\) results in more aggressive rescaling than suggested by
\(p\) alone. Note that when \(p_{\mathrm{eff}} < 2\) we actually invert
the behavior and shift the focus away from the center, in general this
is not recommended.

\hypertarget{sec:prep}{%
\subsection{Data preprocessing}\label{sec:prep}}

Before applying the transformation, it can be important to preprocess
the data, as is generally required with methods for the visualisation of
high-dimensional data. In particular, when working with projections that
combine multiple variables it is important to ensure a common scale
across the dimensions. This is especially important with the sage
display, where the overall scale (set by the maximum radius \(R\))
determines the effect of the transformation. Scaling is commonly also
combined with centering as a preprocessing step. This is less important
in our implementation because we always center the projected data before
applying our transformation (i.e.~the projected mean defines the origin
of our polar coordinate system in the plane). Centering and scaling
together would be achieved by standardizing each variable in the data to
be centered at 0 and have a variance equal to 1.

We may also wish to apply non-linear transformations before visualising
the data or combining variables in a projection. One example would be to
use a log transformation in the case of a variable that has a skewed
distribution. These types of transformations will generally change the
distribution across the high-dimensional volume (and in particular also
the position of the mean), and can thus make the sage display more
useful for a wide range of observed data distributions.

The third common pre-processing would involve a dimension reduction
step. Even with a relatively low number of dimensions, e.g.~with
\(p=10\), we see noticeable piling. However, we are often confronted
with much higher dimensional datasets in which many of the variables are
primarily adding noise. In this case it is advisable to first reduce the
number of dimensions, for example using principal components analysis.
Other dimension reduction techniques could be used. The sage tour would
then be used to view the first few components, which would likely
provide more insight about the data structures than is possible from
only the first two.

\hypertarget{sec:implementation}{%
\subsection{Implementation as a dynamic
display}\label{sec:implementation}}

While the radial transformation can in general be used with any
low-dimensional display that suffers from data crowding, it is most
useful when combined with a dynamic display showing a sequence of
interpolated low-dimensional projections obtained when running a tour.
We have implemented it as a new display method called
\texttt{display\_sage} in the \texttt{tourr} package \citep{tourr} in R
\citep{rref}.

We can think of the display functions as part of a data pipeline
obtained when running a tour. The initial step is pre-processing the
data, given by \(\mathbf{X}\), an \(n \times p\) matrix containing \(n\)
observations in \(p\) dimensions. Typically, this includes centering and
scaling, using either the overall range or the variance. Ensuring a
common scale of all variables, comparable to the selected scale
parameter \(R\), is especially important with the new display. The tour
then iterates over the following steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Obtain projection matrix \(\mathbf{A}\). For \(d\)-dimensional
  projections this is an orthonormal \(p \times d\) matrix. To ensure
  the smooth rotation of projections, each new \(\mathbf{A}\) is
  obtained as an interpolated step in the sequence, as explained in
  \citet{BCAH05}.
\item
  Project the data by computing
  \(\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}\).
\item
  Map \(\mathbf{Y}\) to the display to re-draw the projected data. For
  \(d=2\) this typically maps the projected points onto a scatter plot
  display. With the new display we first transform \(\mathbf{Y}\) as:

  \begin{itemize}
  \tightlist
  \item
    Center the 2-dimensional matrix \(\mathbf{Y}\) and compute its polar
    coordinate representation \((r_y, \theta_y)\).
  \item
    For each observation, first use Eq. \ref{eq:cutR} to get the trimmed
    radius \(r_y^{\mathrm{trim}}\) within the specified range, and then
    apply the radial transformation defined in Eq. \ref{eq:resc} to
    obtain \(r'_y\).
  \item
    Use the transformed radial coordinate \(r'_y\) to re-compute the
    mapping onto Euclidean coordinates \((y'_1, y'_2)\).
  \end{itemize}
\item
  To fit the projected and transformed data \(\mathbf{Y}'\) onto the
  plotting canvas ranging between \([-1,1]\), we rescale each mapped
  observation \(y\) using a scaling parameter \(s\).
\end{enumerate}

The display can be added when calling the \texttt{animate} function in
\texttt{tourr}, as

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tourr}\SpecialCharTok{::}\FunctionTok{animate}\NormalTok{(}
\NormalTok{  data,}
  \AttributeTok{tour\_path =}\NormalTok{ tourr}\SpecialCharTok{::}\FunctionTok{grand\_tour}\NormalTok{(),}
  \AttributeTok{display =} \FunctionTok{display\_sage}\NormalTok{(gam, R, half\_range)}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\noindent and uses \texttt{gam} to set the \(\gamma\) parameter for
computing \(p_{\mathrm{eff}}\), and the overall range \texttt{R} used
for trimming. Both these parameters are described in Section
\ref{sec:params} above.

For consistency with other display functions in the \texttt{tourr}
package we also pass in the scale parameter \(s\) via the
\texttt{half\_range} argument. By default this is evaluated from the
range of the data, but kept as a free parameter to zoom in (without any
transformation) if necessary. This can be important, e.g.~when the range
is large because of an outlying point. Note here that the ratio \(R/s\)
sets the scale for fitting the displayed data on the plotting canvas, by
default \(s = R\). When adjusting \(R\) the user should take care to
adjust \(s\) accordingly.

\hypertarget{sec:application}{%
\section{Applications}\label{sec:application}}

To illustrate the benefit of using the reverse transformation for
examining data using a tour, four applications are shown: clustering of
single cell RNA-seq, classifying hand-sketched images, comparing physics
experiments, and the classical pollen data. The pollen data example is
used to illustrate the effect of parameter choices in the sage tour.

\hypertarget{sec:appl1}{%
\subsection{Clustering single-cell RNA-seq data sets}\label{sec:appl1}}

In the analysis of single cell RNA-seq data, cluster analysis is
performed to detect cell types and characterize the expression of genes
that define those cell types, and the relative orientation of the cell
types to each other (trajectory analysis) \citep{Amezquita2020-at}.
Generally, for cluster verification, analysts use embedding methods like
t-SNE to verify the placement and meaning of clusters from a clustering
algorithm. An alternative is to use a tour on a small number of
principal components to examine the clusters relative to gene
expression.

Here we compare the sage display and regular tour display on mouse
retinal single cell RNA-seq data from \citet{Macosko2015-ot}. The raw
data consists of a 49,300 cells and was downloaded using the using the
\texttt{scRNAseq} Bioconductor package \citep{scRNAseq-d}. We use a
standard workflow for pre-processing and normalizing this data
(described by \citet{Amezquita2020-at}): quality control was performed
using the \texttt{scater} package \citep{McCarthy2017} and
\texttt{scran} \citep{Lun2016} was used to transform and normalize the
expression values and select highly variable genes (HVGs). The top ten
percent of the most HVGs were used as features to subset the normalized
expression matrix and compute the principal components. Using the first
25 PCs we built a shared nearest neighbors graph (with \(k = 10\)) and
clustered this graph using Louvain clustering, resulting in 11 clusters
being formed \citep{Blondel2008-bx}.

A tour is run on the first five PCs (approximately 20\% of the variance
in expression), on a weighted subsample of cells based on their cluster
membership - 4,590 cells. For the sage display, we set \(\gamma = 3\),
fixing the effective dimensionality of the data to
\(p_{\mathrm{eff}} = 15\). The PCs are scaled to have zero mean and unit
variance. Here we focus on comparing three of the clusters. In the PC
plots they look very similar, begging the question whether they should
be considered to be separate groups. Figure \ref{fig:mouse-cluster}
shows selected frames from a default tour (top row) and the sage tour
(bottom row). The columns show the same projection, with the difference
being that the sage transformation is applied in the sage tour
projections. The full animations are available in the supplementary
material. The static plots serve to illustrate the main points, but we
encourage the reader to look at the tour animations to fully appreciate
the advantage of the sage display.

\begin{figure}

{\centering \includegraphics[width=0.9\textwidth]{paper_files/figure-latex/mouse-cluster-1}

}

\caption{Selected frames from using a tour of the mouse data with the default tour display (top), and the sage display with $\gamma=3$ (bottom). Three selected clusters are highlighted in color, all other points are shown in grey. Using the sage display mitigates overplotting and provides a better understanding of cluster separation.}\label{fig:mouse-cluster}
\end{figure}

Using the default tour display (Figure \ref{fig:mouse-cluster}, top),
the three clusters (dark green, blue, and yellow) are obscured by points
in other clusters as we move through the frames of the animation. The
points in the dark green cluster are overlapping those found in the
yellow and blue clusters; and it is difficult to see if there is any
separation between the blue and yellow clusters. In contrast, the sage
display (Figure \ref{fig:mouse-cluster}, bottom), expands the center of
projection, and results in the differences between the three clusters
being more visible. Particularly, the relative positions of the yellow
and blue clusters are easier to see. While these clusters are distinct
from the dark green cluster, in most frames they are still overlapping
and mixed together, providing evidence that it may be appropriate to
consider them a single cluster. Conversely, it can be seen that the dark
green cluster is distinctly separated from the other two in some
projections. The sage tour makes these comparisons a little easier.

\hypertarget{sec:appl2}{%
\subsection{Classifying hand-sketches}\label{sec:appl2}}

We next use the new display to look at different distributions of images
from the Google QuickDraw collection \citep{quickdraw-paper}. These are
\(28\times28=784\) pixel grey scale data that are available publicly. In
this example, we sample 1000 images from three types of sketches
(banana, cactus, crab) and see if we can separate the classes in the
high-dimensional parameter space.

We reduce the dimensionality from 784 variables to the first 5 PCs,
which captures approximately 20 per cent of the variation of the data.
Before applying the tour we rescale each component to have mean zero and
unit variance. To account for the dimension reduction before
visualization we set \(\gamma=2\) for the sage display.

\begin{figure}

{\centering \includegraphics[width=0.9\textwidth]{paper_files/figure-latex/sketches-1}

}

\caption{Selected frames  of the tour run on the sketches data using the default tour display (top), and using the sage display with $\gamma=2$ (bottom). Three types of sketches are indicated by color: banana (green), cactus (orange) and crab (purple). Overplotting of points is a problem for the grand tour display, while the sage display reveals low density near the center.}\label{fig:sketches}
\end{figure}

Figure \ref{fig:sketches} shows the grand tour on the PCs, where green
points correspond to the banana class, orange points represent the
cactus class and purple points are the crab class. In the selected
frames of both displays points belonging to the cactus class are
concentrated near the center, however on the default display (Figure
\ref{fig:sketches}, top) there is overplotting: points from other
classes overlap those in the cactus class. The sage display (Figure
\ref{fig:sketches}, bottom) helps reduce overplotting, it is easier to
see that the centers of class are separated and that there is
substructure in the banana class, which further collapses into two
subgroups.

The animated sage tour available in the supplementary material further
reveals a low density of points near the center of the distribution:
observing the movement of points when rotating the viewing angle shows
that even the cactus class is clustering away from the mean.

\hypertarget{sec:appl3}{%
\subsection{Comparing physics experiments: PDFSense}\label{sec:appl3}}

Data were obtained from CT14HERA2 parton distribution function fits and
describe the sensitivity of fit parameters to experimental measurements
\citep{Wang:2018heo}. There are 28 parameters, and varying one at a time
to move \(\pm 1 \sigma\) away from the `best fit point' (maximum
likelihood estimate) provides our input variables, labeled X1-X56. Each
of the 2808 observations corresponds to a physical observable and
measures how the fit prediction changes along the 56 directions in
parameter space. Points are grouped based on the underlying process in
the experiment, which is mapped to color in the following. With the
analysis of the distribution along these variables X1-X56 we can
understand to what extent each experimental measurement provides new
information for the global fit. For example, orthogonality between
groups marks complementary constraints, and outlying points are
considered as important for future fits, see discussion in
\citet{Cook:2018mvr}.

\begin{figure}

{\centering \includegraphics[width=0.9\textwidth]{paper_files/figure-latex/pdfsense-1}

}

\caption{Selected frames of the tour of the pdfsense data using the default tour display (top), and using the sage display with $R=10$ (bottom). Different underlying physical processes are shown by color and we can see orthogonality between the three groups. The sage display preserves the overall structure while revealing details that are hidden near the center in the default display.}\label{fig:pdfsense}
\end{figure}

Following the processing described there, we tour the first 6 PCs,
rescaled to have zero mean and unit variance. In Figure
\ref{fig:pdfsense} we see that the sage display with \(R = 10\) (Figure
\ref{fig:pdfsense}, bottom), maintains the overall shape of the data
seen using the default tour display (Figure \ref{fig:pdfsense}, top).
The different physical process, shown in different colors, are indeed
orthogonal in the parameter space, as can be seen most clearly by
looking at the animations available in the supplementary material.

The particular structure of this distribution, with some clusters
extending linearly away from the center and a set of outlying points,
results in poor use of the plotting space, and high level of clustering
near the center. For example, focusing on the blue cluster, we can see
that it extends out along different directions, but it can be
challenging to observe how the points move under the tour rotation, as
overplotting becomes an issue when points move through the center. Here,
the new display (bottom row) shows a clearer view.

\hypertarget{sec:appl4}{%
\subsection{Tuning the parameters: Pollen}\label{sec:appl4}}

The classical pollen data is useful to demonstrate the trimming and
tuning parameters. The five-dimensional data set was simulated by David
Coleman of RCA Labs, for the Joint Statistics Meetings 1986 Data Expo
\citep{pollen}, and is an example of a hidden structure near the center
of a distribution. The data are standardized by centering and scaling
such that the standard deviation of each variable is equal to one.

Neither the standard tour display nor the sage display with default
settings (\(R=6.6\) which is set by the data scale, and \(\gamma=1\))
reveals the structure (left plot in Figure \ref{fig:pollen}). We can use
either \(\gamma\), \(R\) or a combination of the two to zoom in further
near the center. For example, we can use trimming \((R=1, \gamma=1)\)
(middle plot) or tuning \((R=6.6,\gamma=20)\) (right plot) as shown in
Figure \ref{fig:pollen}. There is an approximate equivalence between the
results obtained using either tuning or trimming, and both views clearly
reveal the word ``EUREKA'' hidden in the distribution.

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{../pngs/pollen_sage-088.png}
\includegraphics[width=0.3\textwidth]{../pngs/pollen_sage_R1-088.png}
\includegraphics[width=0.3\textwidth]{../pngs/pollen_sage_gam20-088.png}
\caption{Selected views of the pollen data in the new sage display, with default settings (left), setting $R=1$ (middle) and $\gamma=20$ (right). We can tune either $\gamma$, $R$ or a combination of the two to reveal the word "EUREKA" near the center of the distribution.}
\label{fig:pollen}
\end{figure}

While the static views look very similar, comparing the tour animations
(available in the supplementary material) reveals some differences
between the display with trimming or tuning. When trimming (by setting
\(R=1\)) the focus is clearly on the center of the distribution, and
most points get pushed out towards a maximum radius circle. On the other
hand, tuning the display by setting \(\gamma=20\) preserves the
elliptical shape of the distribution, making it easier to see
correlation patterns.

\hypertarget{sec:concl}{%
\section{Discussion}\label{sec:concl}}

This paper has introduced the sage tour, which reduces the data crowding
effects that occur when taking low-dimensional projections of
high-dimensional data. This new technique is easily incorporated into
exploratory high-dimensional data analysis, and applications shown in
Section \ref{sec:application} provide examples of the following tasks:

\begin{itemize}
\tightlist
\item
  clustering: the sage display uncovered clusters that were originally
  obscured by data piling, while still giving the viewer an accurate
  assessment of the size of a cluster, and their relative orientation,
  as shown in the single cell RNA-seq example (Section \ref{sec:appl1})
\item
  classification: the sage display decreases the number of overlapping
  points between classes and provides better visual separation between
  classes compared to the regular tour, as shown in the sketches example
  (Section \ref{sec:appl2})
\item
  shape analysis: the sage display helps us understand structures across
  multiple dimensions, for example orthogonality between multiple
  groups, as shown in the pdfsense example (Section \ref{sec:appl3})
\item
  needle discovery: the sage display allows to find hidden signal that
  is concealed by the density of points around the center of the
  projection, as shown in the pollen example (Section \ref{sec:appl4})
\end{itemize}

The approach provides interpretable visualization that captures high
dimensional information and preserves global structure, and it is
complementary to non-linear dimension reduction techniques. For example,
when visualizing clusters, the sage display enables an assessment of
cluster shapes, and accurately captures relative position and
orientation. The burning sage transformation is global and does not
magnify local structure like t-SNE does.

\textcolor{red}{Note here that adjusting the resolution using the burning sage transformation will lead to an overall decrease in densitiy near the center. When studying the separation between groups it is important to keep in mind that the effect is the same within and between clusters. Thus, if points within a cluster remain grouped while separation between clusters becomes more apparant, this can be interpreted as a true finding. Conversely, in the case of a visual artifact we expect a general reduction of density that does not lead to better separation between the clusters.}

An alternative is the slice tour \citep{sliceTour} which allows
distributions of points around the center of the data to be explored
using sections instead of projections. Instead of adjusting for effects
from projecting a high-dimensional volume, the solution with the slice
tour is to only highlight projected points if they are within a small
volume around the current projection plane, thus showing a local view of
the distribution. The slice tour is useful when there are large numbers
of observations or if there is concave structure in the data. Note that
only few methods are currently available for looking at sections of the
data in a systematic fashion, and further work is needed to fully
explore the potential of this type of visualisation. In smaller datasets
we may also use interactive approaches such as linked brushing to
explore for local structures or conditional features.

The tuning parameters can be used to more aggressively expand the center
of the display. All of the examples shown had some tuning. The last
example demonstrated how points away from the projected center get moved
to the edge of the hypersphere as \(\gamma\) is increased or \(R\) is
decreased. With more center magnification, the non-linear transformation
can introduce distortions, but this is a well known problem for any
non-linear dimension reduction technique including t-SNE. However,
unlike t-SNE, any distortion introduced by the sage display is
interpretable because it is controlled by a simple function (Eq.
\ref{eq:resc}).

The sage display is fast to compute, which lends itself to being
embedded into an interactive interface. An ideal interface would allow
real time changes to the parameters of the transformation. This would be
especially useful when coupled with linked brushing in complementary
views.

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

The authors gratefully acknowledge the support of the Australian
Research Council. The paper was written in \texttt{rmarkdown}
\citep{rmarkdown} using \texttt{knitr} \citep{knitr}.

\hypertarget{supplementary-material}{%
\section*{Supplementary material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary material}

The source material for this paper is available at
\url{https://github.com/uschiLaa/burning-sage}. The animated gifs for
all applications are also included in html files in the supplementary
material.

\bibliographystyle{tfcad}
\bibliography{biblio.bib}

\end{document}